Inverse Problems --- Example Sheet 1

### 2020-11-01 inverse_problems/inverse_ES1_handin.tex
yolo

	\item 	In example 2.2.12, it is shown that for this operator, the Picard criterion is
	\begin{equation} \label{eq:picard}
	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2,
	\end{equation}
	where $\sigma_j = \frac2{(2j - 1)\pi}$.

@@

\begin{question}
	\begin{enumerate}[(a)]
		\item

	Let $m, n \in \NN$ with $m \geq n \geq 2$. Compute the Moore-Penrose inverses of the following matrices:
	\begin{enumerate}[(i)]
		\item $A = (1, 1, \dotsc, 1) \in \RR^{1 \times n}$;
		\item $A = \diag(a_1, \dotsc, a_n) \in \RR^{n \times n}$ with $a_j \in \RR$ for $j = 1, \dotsc, n$;
		\item $A \in \RR^{m \times n}$ with $A\T A = I_n$. 
	\end{enumerate}

\item Let $a, b \in \RR$ with $a < b$. Compute the Moore-Penrose inverse of the operator $A \colon L^2([a, b]) \to \RR$ with $Au = \int_a^b u(x) \dd{x}$. 
	\end{enumerate}
\end{question}

\begin{solution}
	\begin{enumerate}[(a)]
		\item \begin{enumerate}[(i)]
			\item Clearly $\Ran(A) = \RR$ and $\Nul(A)\P = \Span\qty{\vec e}$ where $\vec e$ is the all-ones vector. So $A\D$ must map $x \in \RR$ to $\vec e/n$, and therefore we have
			\[
			A\D = \vec e/n \in \RR^{n \times 1}.
			\]

			\item Clearly we have $\Ran(A) = \Span\qty{\vec e_j \mid a_j \neq 0} = \Nul(A)\P$ while $\Ran(A)\P = \Span\qty{\vec e_j \mid a_j = 0} = \Nul(A)$.

			It is easily seen that
			\[
			A\D = \diag(b_1, \dotsc, b_n) \in \RR^{n \times n}, \quad\text{where } b_i = \begin{cases}
				a_i^{-1} &\text{if $a_i \neq 0$}; \\
				0 &\text{if $a_i = 0$}. 
			\end{cases}
			\]

			\item 
%			If $A\T A = I$, then the columns $\vec u_1, \dotsc, \vec u_n$ of $A$ are orthogonal. In particular, $A$ is injective (so $\Nul(A) = \qty{0}$) and $\Ran(A) = \Span\qty{\vec u_1, \dotsc, \vec u_n}$. Extend $\vec u_1, \dotsc, \vec u_n$ to an orthonormal basis $\vec u_1, \dotsc, \vec u_m$ of $\RR^m$, then we must have
%			\[
%			A\D\qty(\sum_{i=1}^m \alpha_i \vec u_i) = \mqty(\alpha_1 \\ \vdots \\ \alpha_n) \in \RR^n.
%			\]
%			From this expression it is easily seen that $A\D = A\T$.
			We claim $A\D = A\T$. For this, we check the first two Moore-Penrose equations: for the first one, we have $A\T A = I_n = P_{\RR^n} = P_{\Nul(A)\P}$ since $\Nul(A) = \qty{0}$.  
			For the second one, let $\vec u_1, \dotsc, \vec u_n \in \RR^m$ be the columns of $A$ (which are orthonormal), then we have
			\[
			AA\T = \sum_{i=1}^n \vec u_i \vec u_i\T = P_{\Span\qty{\vec u_1, \dotsc, \vec u_n}} = P_{\clos{\Ran(A)}}. 
			\]
		\end{enumerate}

	\item Clearly we have $\Ran(A) = \RR$ while $\Nul(A) = \qty{u \mid \int_a^b u(x) \dd{x} = 0}$. It is also easily seen that 
	\[
	\Nul(A)\P = \qty{v \mid \int_a^b v(x) u(x) \dd{x} = 0 \text{ if $\int_a^b u(x) \dd{x} = 0$}} = \Span\qty{1}. 
	\]

	Therefore we simply have that $A\D$ maps a constant $c \in \RR$ to the constant function $\frac{c}{b-a}$. 
	\end{enumerate}
\end{solution}

\begin{question}
	Many forward problems are either modelled as convolutions or they are modelled as the composition of several components, one of which is a convolution. Therefore convolutions play an important role in inverse problems. As in Exercise 1, let $\Omega =[0, 1]^2$ be the unit square and let $\XX = L^2(\Omega)$. A convolution is the special case of an integral operator $A \colon \XX \to \XX$ where the kernel has a simple structure: 
	\[
	(Au)(y) \ceq \int_\Omega k(y - x) u(x) \dd{x},
	\]
	for $k \in L^2(\Omega)$. It follows easily from Exercise 1 that $A$ is linear and bounded. 
	\begin{enumerate}[(a)]
		\item Although shown in general in Exercise 1, give an explicit form for the adjoint of the convolution. 
		\item Let $f = Au$. It follows from the convolution theorem that a convolution can be inverted by means of the Fourier transform
		\begin{equation} \label{eq:fourier}
		u = (2\pi)^{-\frac n2} \FF^{-1} \qty(\frac{\FF(f)}{\FF(k)}),
		\end{equation}
		where $\FF$ is the Fourier transform and $\FF^{-1}$ its inverse. Implement this formula in \emph{MATLAB} to deblur the blurry tree image $f$ generated by the script \verb|ex4b_generate_data.m|. Note that the script also outputs $\FF(k)$. Add some noise to the data and show that the inversion formula is ill-conditioned.

		Hint: Make use of the \emph{MATLAB} commands \verb|fft2| and \verb|ifft2|.

		\item Reformulate \cref{eq:fourier} so that the denominator is non-negative and give a stable approximation of this formula. Implement this formula in MATLAB and empirically show that it is stable.

		Hint: Make use of the \emph{MATLAB} command \verb|conj|. 
	\end{enumerate}
\end{question}

\begin{proof}
	\begin{enumerate}[(a)]
		\item We have by question 1 that
		\[
		(A^* v)(y) = \int_\Omega k(x - y)v(x) \dd{x}.
		\]
		\item See matlab code. 
		\item We can write $\FF(f) / \FF(k) = \FF(f) \overline{\FF(k)} / \abs{\FF(k)}^2$. 
	\end{enumerate}
\end{proof}
\end{document}

>>>>

	\item 	In example 2.2.12, it is shown that for this operator, the Picard criterion is
	\begin{equation} \label{eq:picard}
	2\sum_{j=1}^\infty  \sigma_j^{-2} \qty(\int_0^1 f(s) \sin(\sigma_j^{-1} s) \dd{s})^2 < \infty,
	\end{equation}
	where $\sigma_j = \frac2{(2j - 1)\pi}$.

@@

\end{document}
